Covariate analysis of MSSM reprocessed counts (with CQN normalisation and covariates adjusted)

### 2017-04-24 code/Rmd/MAYO_REPROCESSED.Rmd
Fixed outlier expression/ winsorised MSSM data

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_CBE_TCX_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MAYO_CBE_TCX_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_CBE_TCX_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MAYO_CBE_TCX_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_CBE_TCX_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('MAYO_CBE_TCX_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

>>>>

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts.NA %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_CBE_TCX_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MAYO_CBE_TCX_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
tmp = CQN.GENE_EXPRESSION$y
tmp[log.mat] = NA
tmp %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_CBE_TCX_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MAYO_CBE_TCX_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
tmp = CQN.GENE_EXPRESSION$offset
tmp[log.mat] = NA
tmp %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_CBE_TCX_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('MAYO_CBE_TCX_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

### 2017-04-24 code/Rmd/MSSM_REPROCESSED.Rmd
Fixed outlier expression/ winsorised MSSM data

### Conditional Quantile Normalisation (CQN)
Initial normalisation usign cqn and dispersion control with voom (with NULL design)
```{r initial.voom.normalisation, fig.height=5, fig.width=5}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[rownames(PROCESSED_COUNTS$filteredExprMatrix), 'percentage_gc_content'],
                          lengths = GENE.LEN[rownames(PROCESSED_COUNTS$filteredExprMatrix), 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)

# Set gene counts in specific samples that are deviating 3 sd from other samples to 0
log.mat = apply(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  return((x < (mn-3*std.dev)) | (x > (mn+3*std.dev)))
}) %>% t
PROCESSED_COUNTS$filteredExprMatrix$counts[log.mat] = 0

# Re-compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[rownames(PROCESSED_COUNTS$filteredExprMatrix), 'percentage_gc_content'],
                          lengths = GENE.LEN[rownames(PROCESSED_COUNTS$filteredExprMatrix), 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
```

```{r explore.data, cache=FALSE, fig.height=5, fig.width=10}
# Plot abberent distribution of logcpm counts
tmp1 = (CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset) %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.')

p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(BrainRegion+.~Diagnosis, scale = 'free')
p
```
Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
### Decomposition of expression
Clustering of library normalised unadjusted data (with NULL design)
```{r decompse.normalise.data1.1, fig.height=5, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset, scale.=T, center = T)

@@

  # Calculate correlation values of random effects
  VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$individualIdentifier)

@@

  # Fit linear model using new weights and new design
  VOOM.ADJUSTED.FIT = lmFit(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset,
                            design = DM1$design,
                            weights = VOOM.GENE_EXPRESSION$weights,
                            block = COVARIATES$individualIdentifier, 
                            correlation = correlation$cor)

  # Residuals after normalisation
  RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset)

@@

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

@@

tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`
### Sanity check
```{r residual.adj, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)

residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2.1, fig.height=6, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

@@

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp, fig.height=6, fig.width=15}
# Get design matrix
COVARIATES.tmp = COVARIATES %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  dplyr::mutate(BrainRegion = factor(BrainRegion))
DESIGN = getDesignMatrix(COVARIATES.tmp[, c('BrainRegion', postAdjustCovars), drop = F], Intercept = F)
ind = grep('BrainRegion', colnames(DESIGN$design))
DESIGN$design[,ind] = DESIGN$design[,ind] * as.numeric(COVARIATES.tmp[,'CDR'])
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

@@

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
contrast = makeContrasts(contrasts = c("BrainRegionFP", "BrainRegionIFG", "BrainRegionPHG", "BrainRegionSTG"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

@@

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
cntr = c()
for (region in c('FP', 'STG', 'PHG', 'IFG')){
  tmp = combn(paste0('BrainRegion.Diagnosis', region, '.', c('OTHER','AD','CONTROL')),2)
  cntr = c(cntr, apply(tmp, 2, paste, collapse = '-'))
}

@@

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=15, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE %>% dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = log10(gene.length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(gene.length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(BrainRegion~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')

pl[[2]] = ggplot(DE %>% dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = percentage_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(BrainRegion~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')

pl[[3]] = ggplot(DE %>% dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(BrainRegion~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')

@@

# Estimate surrogate variables
SURR.VAR = sva::sva(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset,
                    MODEL1, MODEL0, n.sv = NUM.SV, B = 30)$sv

@@

# Calculate correlation values of random effects
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset
correlation = parallelDuplicateCorrelation(VOOM.SVA.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Calculate correlation values of random effects
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset
correlation = parallelDuplicateCorrelation(VOOM.NET.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSSM_FP_STG_PHG_IFG_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MSSM_FP_STG_PHG_IFG_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSSM_FP_STG_PHG_IFG_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MSSM_FP_STG_PHG_IFG_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSSM_FP_STG_PHG_IFG_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('MSSM_FP_STG_PHG_IFG_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

>>>>

### Conditional Quantile Normalisation (CQN)
Initial normalisation usign cqn and dispersion control with voom (with NULL design)
```{r initial.voom.normalisation, fig.height=5, fig.width=5}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[rownames(PROCESSED_COUNTS$filteredExprMatrix), 'percentage_gc_content'],
                          lengths = GENE.LEN[rownames(PROCESSED_COUNTS$filteredExprMatrix), 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E.NA = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset

# Set gene counts in specific samples that are deviating 3 sd from other samples to 0
log.mat = apply(CQN.GENE_EXPRESSION$y+CQN.GENE_EXPRESSION$offset, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)
  return((x < (mn-3*std.dev)) | (x > (mn+3*std.dev)))
}) %>% t
PROCESSED_COUNTS$filteredExprMatrix$counts.NA = PROCESSED_COUNTS$filteredExprMatrix$counts 
PROCESSED_COUNTS$filteredExprMatrix$counts.NA[log.mat] = NA
PROCESSED_COUNTS$filteredExprMatrix$counts[log.mat] = 0

CQN.GENE_EXPRESSION$E.NA[log.mat] = NA
CQN.GENE_EXPRESSION$E.zero = CQN.GENE_EXPRESSION$E.NA
CQN.GENE_EXPRESSION$E.zero[log.mat] = 0
```

```{r explore.data, cache=FALSE, fig.height=5, fig.width=10}
# Plot abberent distribution of logcpm counts
tmp1 = (CQN.GENE_EXPRESSION$E.NA) %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.')

p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(BrainRegion+.~Diagnosis, scale = 'free')
p
```
Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(CQN.GENE_EXPRESSION$E.zero))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
### Decomposition of expression
Clustering of library normalised unadjusted data (with NULL design)
```{r decompse.normalise.data1.1, fig.height=5, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.zero, scale.=T, center = T)

@@

  # Calculate correlation values of random effects
  VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E.NA
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$individualIdentifier)

@@

  # Fit linear model using new weights and new design
  VOOM.ADJUSTED.FIT = lmFit(CQN.GENE_EXPRESSION$E.NA,
                            design = DM1$design,
                            weights = VOOM.GENE_EXPRESSION$weights,
                            block = COVARIATES$individualIdentifier, 
                            correlation = correlation$cor)

  # Residuals after normalisation
  RESIDUAL.GENE_EXPRESSION = list()
  RESIDUAL.GENE_EXPRESSION$expr.NA = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                        CQN.GENE_EXPRESSION$E.NA)
  RESIDUAL.GENE_EXPRESSION$expr.zero = RESIDUAL.GENE_EXPRESSION$expr.NA
  RESIDUAL.GENE_EXPRESSION$expr.zero[is.na(RESIDUAL.GENE_EXPRESSION$expr.zero)] = 0

@@

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION$E.zero, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

@@

tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`
### Sanity check
```{r residual.adj, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION$expr.zero, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)

residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION$expr.zero))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2.1, fig.height=6, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION$expr.zero, scale.=T, center = T)

@@

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp, fig.height=6, fig.width=15}
# Get design matrix
COVARIATES.tmp = COVARIATES %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  dplyr::mutate(BrainRegion = factor(BrainRegion))
DESIGN = getDesignMatrix(COVARIATES.tmp[, c('BrainRegion', postAdjustCovars), drop = F], Intercept = F)
ind = grep('BrainRegion', colnames(DESIGN$design))
DESIGN$design[,ind] = DESIGN$design[,ind] * as.numeric(COVARIATES.tmp[,'CDR'])
colnames(DESIGN$design)[ind] = paste0(colnames(DESIGN$design),'.CDR')
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

@@

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E.NA
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E.NA
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
contrast = makeContrasts(contrasts = c("BrainRegionFP.CDR", 
                                       "BrainRegionIFG.CDR", 
                                       "BrainRegionPHG.CDR", 
                                       "BrainRegionSTG.CDR"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

@@

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E.NA
FIT = lmFit(VOOM.WEIGHTS)

# Fit contrast
cntr = c()
for (region in c('FP', 'STG', 'PHG', 'IFG')){
  tmp = combn(paste0('BrainRegion.Diagnosis', region, '.', c('CONTROL', 'AD', 'OTHER')),2)
  cntr = c(cntr, apply(tmp, 2, paste, collapse = '-'))
}

@@

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=15, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE %>% dplyr::filter(Comparison == 'CONTROL-AD'), 
                 aes(x = log10(gene.length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(gene.length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(BrainRegion~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')

pl[[2]] = ggplot(DE %>% dplyr::filter(Comparison == 'CONTROL-AD'), 
                 aes(x = percentage_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(BrainRegion~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')

pl[[3]] = ggplot(DE %>% dplyr::filter(Comparison == 'CONTROL-AD'), 
                 aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(BrainRegion~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')

@@

# Estimate surrogate variables
SURR.VAR = sva::sva(CQN.GENE_EXPRESSION$E.zero,
                    MODEL1, MODEL0, n.sv = NUM.SV, B = 30)$sv

@@

# Calculate correlation values of random effects
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E.NA
correlation = parallelDuplicateCorrelation(VOOM.SVA.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Calculate correlation values of random effects
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E.NA
correlation = parallelDuplicateCorrelation(VOOM.NET.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts.NA %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSSM_FP_STG_PHG_IFG_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MSSM_FP_STG_PHG_IFG_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
tmp = CQN.GENE_EXPRESSION$y
tmp[log.mat] = NA
tmp %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSSM_FP_STG_PHG_IFG_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MSSM_FP_STG_PHG_IFG_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
tmp = CQN.GENE_EXPRESSION$offset
tmp[log.mat] = NA
tmp %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSSM_FP_STG_PHG_IFG_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('MSSM_FP_STG_PHG_IFG_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
